---
title: "MDS datasets"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    warning = FALSE,
    message = FALSE
)
```

## Transcriptome sequencing of myelodysplasia 
| Dataset         | ID       | Technology | Samples |
| --------------- | -------- | ---------- | ------- |
| [EGAD00001003891](https://ega-archive.org/datasets/EGAD00001003891) | Illumina | HiSeq 2500 | 266     |

### Dataset Description
Transcriptome sequencing was performed on 214 patients with myelodysplasia in this study. RNA was obtained from bone marrow CD34+ cells (n=100) and/or bone marrow mononuclear cells (n=165). Transcriptome sequencing was performed for both cell fractions in 51 patients. A total of 211 patients were genotyped by targeted deep sequencing. For controls, bone marrow CD34+ cells and bone marrow mononuclear cells were obtained from three healthy adults each.

### Usage
Data were downloaded from EGA with approval and must not be used for any purpose other than PSON studies. 

Shiozawa et al 2017 *Blood* [![DOI](https://zenodo.org/badge/DOI/10.1182/blood-2017-05-783050.svg)](http://doi.org/10.1182/blood-2017-05-783050)

Shiozawa et al 2018 *Nat Comm* [![DOI](https://zenodo.org/badge/DOI/10.1038/s41467-018-06063-x.svg)](http://doi.org/10.1038/s41467-018-06063-x)

```{r setup}
library(haemdata)
use_pinboard("onedrive")

se <- get_pin("hsa_mrna_mds_GENCODEr40_qc.rds")

pca <- pca_se(se, col_by = "tissue")
```

```{r pca_plots}
pca$biplot
pca$pairs_plot
```
```{r pca_corr, fig.height=9, fig.width=8}
pca$correlation_plot
```
